A Constraint Solving Approach to Parikh Images of Regular Languages

### 2023-07-11 preliminaries.tex
polishing

We define the image of this map, the \textit{Parikh image}, of some subset of
the language $\Language \subseteq \Strings$ as:
\[
\ParikhMap(\Language) = \Set{\ParikhMap(x) \SuchThat x \in \Language}
\]

Thus, we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1,
1\right], \left[1, 2\right]\right\}$. We also sometimes use the standard
notation $\#l(w)$ to talk about an individual letter $l$ in a word $w$. For
example, for the Parikh vector above, we would have $\CountOf{a} = 1$.

@@

Since Parikh images are semilinear, any Parikh image can be written as a set of
linear equations. The version for context-free grammars presented in
\cite{generate-parikh-image} can be straightforwardly adapted for use with a
product of regular languages or Parikh automta. For an intuition, the approach
consists of first computing the product, then assigning each state and
transition an existentially quantified non-negative integer variable, then
describing all paths through the automaton through two sets of equations: one
that preserves the incoming flow across the variables of a state's incoming
transitions is equal to the flow across its outgoing transitions, and one set of
constraints that enforces coherence in the presence of loops by ordering states
by distance in a spanning tree rooted in the initial state.

>>>>

We define the image of this map, the \textit{Parikh image}, of some
language $\Language \subseteq \Strings$ as:
\[
\ParikhMap(\Language) = \Set{\ParikhMap(x) \SuchThat x \in \Language}
\]

Thus, we would have $\ParikhMap(\left\{ab, abb\right\}) = \left\{\left[1,
1\right], \left[1, 2\right]\right\}$. We also sometimes use the standard
notation $\CountOf{l}(w)$ to talk about an individual letter $l$ in a word $w$. For
example, for the Parikh vector above, we would have $\CountOf{a}(abb) = 1$.

@@

It is known that the Parikh image of any context-free language can be
described by a linear-size existential Presburger
formula~\cite{generate-parikh-image}. This representation can be
straightforwardly adapted for use with a product of regular
languages. For an intuition, the approach consists of first computing
the product, then assigning each state and transition an existentially
quantified non-negative integer variable, and then describing all
paths through the automaton through two sets of constraints: flow
equations relating the inflow and outflow of each automaton state, and
constraints that enforce connectedness by ordering states by distance
in a spanning tree rooted in the initial state.
